---
title: "Time series decomposition and Forecasting"
output:
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```

```{r, echo = FALSE}
library(fpp3)
library(tidyverse)
library(EpiEstim)
```

```{r, echo=FALSE}
read_csv("covidlive_data_2022-10-02.csv") %>%
  select(-...1) %>% 
  filter(date_confirmation <= '2022-09-09') -> covidlive_ll

covidlive_ll %>%
  group_by(date_confirmation) %>%
  summarise(daily_case = sum(daily_notification)) -> full_data

colnames(full_data) <- c("Date", 'I')
```

# Original data

```{r}
full_data %>%
  as_tsibble(index = Date) -> data_ts

data_ts %>%
  head()
```

```{r}
data_ts %>%
  autoplot()
```

# Decomposition

I use the Multiple seasonal decomposition by Loess, which is STL decomposition. I use the whole sample, the date from 2020-01-25 to 2022-09-09

```{r}
data_ts %>%
  model(stl = STL(I ~ season(period = "week"))) -> decomp_full
```

```{r}
decomp_full %>%
  components() %>%
  autoplot()
```

```{r}
components(decomp_full) %>%
  ggplot() +
  geom_line(aes(x = Date, y = I), color = "grey") +
  geom_line(aes(x = Date, y = season_adjust, color = "season adjust"))
```

```{r}
components(decomp_full) %>%
  filter(Date >= ymd("2022-01-01")) %>%
  ggplot() +
  geom_line(aes(x = Date, y = I), color = "grey") +
  geom_line(aes(x = Date, y = season_adjust, color = "season adjust"))
```

# Data transformation

The season-adjust line, which is red line, looks good.
However, the decomposition plot shows that the season pattern is not consistent. 
The logarithm transformation is applied here. 
The data set include 0, but the domain of logarithm is all number bigger than 0.
Therefore, I remove all data before begining 

```{r}
data_ts %>%
  filter(Date >= ymd("2021-01-01")) %>%
  autoplot(log(I))
```

The **log(I)** looks change rapidly from 2021-01-01 to 2021-07-01, I use data from 2021-08-01 to 2022-09-09

```{r}
data_ts %>%
  filter(Date >= ymd("2021-08-01")) %>%
  autoplot(log(I))
```

The data looks stable.

```{r}
data_ts %>%
  filter(Date >= ymd("2021-07-01")) %>%
  model(stl = STL(log(I) ~ season(period = "week", window = 28))) -> decomp

components(decomp) %>%
  autoplot()
```

The seasonal pattern looks stable after 2022-01-01. 

```{r}
data_ts %>%
  filter(Date >= ymd("2021-07-01")) -> df
components(decomp) -> decomp_comp

ggplot() +
  geom_line(aes(x= Date, y = log(I)), color = "grey", data = df) +
  geom_line(aes(x = Date, y = season_adjust, color = "season_adjust"), data = decomp_comp) +
  ylab("log(I)")
```

The season-adjust line looks very good, it capture the trend of the actual data

```{r}
decomp_comp %>%
  select(Date, season_adjust) -> full_data_adj
```

# Data Frame

```{r}
colnames(full_data_adj) <- c("Date", "I")

data.frame(full_data_adj) -> data
```

# Function

```{r}
r_estiamte <- function(df, start, end, mean, std){
  output <- estimate_R(df, method = "parametric_si",
                  config = make_config(list(mean_si = mean, std_si = std,
                                       t_start = start,
                                       t_end = end)))
  output$dates <- data$Date
  
  return(output)
}

find_r <- function(date, r_df, full_df){
  r_matrix <- matrix(NA, nrow = length(date))
  for (i in 1:length(date)) {
    date_index <-which(full_df$Date == date[i])
    r_index <- which(r_df$R$t_end == date_index)
    r_matrix[i,] <- r_df$R$`Mean(R)`[r_index]
  }
  date_r = tibble("Date" = range, "R" = r_matrix[,1])
  return(date_r)
}

forecast_i <- function(r_date, full_df, r_df){
  
  output_df <- tibble("Date"= as.Date(NA), 
                      "I" = as.numeric(NA), 
                      "Week" = as.numeric(NA))
  
  for (i in 1:dim(r_date)[1]) {
    
    I_renew<-full_df$I[which(full_df$Date <= r_date$Date[i])]
    I_lambda <-I_renew[(length(I_renew) - 99):length(I_renew)]
    data.frame(r_df$si_distr)[,1][1:100] -> si
    predict_w1 <- matrix(NA, nrow = 7, ncol = 1)
    
    for (j in 1:7) {
      element <-  overall_infectivity(I_lambda, si)[100+j-1] * r_date$R[i]
      predict_w1[j,1] <- element
      I_lambda <- append(I_lambda, element)
      si <- append(si, 0)
      
      temp <- tibble("Date" = seq(ymd(r_date$Date[i]), ymd(r_date$Date[i])+6, "day"),
                     "I" = predict_w1[,1],
                     "Week" = i)
    }
    
    output_df <- bind_rows(output_df, temp)
    
  }
  
  output_df %>%
    drop_na() %>%
    mutate(Week = paste0("Period", Week)) -> output_df
  
  return(output_df)
  
}


add_season_pattern <- function(f_data, seasona_data){
  
  f_data$Date - 7 -> season_date
  
  seasona_data %>%
    filter(Date %in% season_date) -> seasona_df
  
  f_data$I <- f_data$I + seasona_df$season_week
  return(f_data)
}
```

# windows

The window begins at 2020/09/01

```{r}
#09/01 708
# one week
t_one <- seq(60, nrow(data)-7)
te_one <- t_one + 7
# two week
t_two <- seq(60, nrow(data)-14)
te_two <- t_two + 14

# three week
t_three <- seq(60, nrow(data)-21)
te_three <- t_three + 21

# four week
t_four <- seq(60, nrow(data)-28)
te_four <- t_four + 28

# five week
t_five <- seq(60, nrow(data)-35)
te_five <- t_five + 35
```

# Estimate R

Using SI mean 4.7, std 2.9

```{r}
res_w1 <- r_estiamte(df = data, start = t_one, end = te_one, mean = 4.7, std = 2.9)
res_w2 <- r_estiamte(df = data, start = t_two, end = te_two, mean = 4.7, std = 2.9)
res_w3 <- r_estiamte(df = data, start = t_three, end = te_three, mean = 4.7, std = 2.9)
res_w4 <- r_estiamte(df = data, start = t_four, end = te_four, mean = 4.7, std = 2.9)
res_w5 <- r_estiamte(df = data, start = t_five, end = te_five, mean = 4.7, std = 2.9)
```

```{r}
tibble("Date" = data$Date[te_one],"R" = res_w1$R$`Mean(R)`, "window" = "w1") -> r_w1
tibble("Date" = data$Date[te_two],"R" = res_w2$R$`Mean(R)`, "window" = "w2") -> r_w2
tibble("Date" = data$Date[te_three],"R" = res_w3$R$`Mean(R)`, "window" = "w3") -> r_w3
tibble("Date" = data$Date[te_four],"R" = res_w4$R$`Mean(R)`, "window" = "w4") -> r_w4
tibble("Date" = data$Date[te_five],"R" = res_w5$R$`Mean(R)`, "window" = "w5") -> r_w5
```

```{r}
bind_rows(r_w1, r_w2) %>%
  bind_rows(r_w3) %>%
  bind_rows(r_w4) %>%
  bind_rows(r_w5) -> df_f
```

```{r}
df_f %>%
  ggplot() +
  geom_line(aes(x = Date, y = R, color = window)) +
  geom_hline(aes(yintercept = 1), color = "grey") +
  theme_bw()
```

The R does not show much difference with the R of original I, which does not do seasonal adjust.

# Forecasting

```{r}
start_date <- ymd("2022-07-01")
end_date <- ymd("2022-09-09")
range <- seq(start_date, end_date,"week")

window_w1 <- find_r(date = range, r_df = res_w1, full_df = full_data_adj)
window_w2 <- find_r(date = range, r_df = res_w2, full_df = full_data_adj)
window_w3 <- find_r(date = range, r_df = res_w3, full_df = full_data_adj)
window_w4 <- find_r(date = range, r_df = res_w4, full_df = full_data_adj)
window_w5 <- find_r(date = range, r_df = res_w5, full_df = full_data_adj)
```

```{r}
window1_predict <- forecast_i(r_date = window_w1, full_df = full_data_adj, r_df = res_w1)
window2_predict <- forecast_i(r_date = window_w2, full_df = full_data_adj, r_df = res_w2)
window3_predict <- forecast_i(r_date = window_w3, full_df = full_data_adj, r_df = res_w3)
window4_predict <- forecast_i(r_date = window_w4, full_df = full_data_adj, r_df = res_w4)
window5_predict <- forecast_i(r_date = window_w5, full_df = full_data_adj, r_df = res_w5)
```

# Add season pattern

```{r}
window1_predict <- add_season_pattern(f_data = window1_predict, seasona_data = decomp_comp)
window2_predict <- add_season_pattern(f_data = window2_predict, seasona_data = decomp_comp)
window3_predict <- add_season_pattern(f_data = window3_predict, seasona_data = decomp_comp)
window4_predict <- add_season_pattern(f_data = window4_predict, seasona_data = decomp_comp)
window5_predict <- add_season_pattern(f_data = window5_predict, seasona_data = decomp_comp)
```

# Graph

```{r}
decomp_comp %>%
  filter(Date >= "2022-07-01") %>%
  ggplot() +
  geom_line(aes(x = Date, y = `log(I)`,color = "Actual"),color = "grey") +
  geom_point(aes(x = Date, y = I,color = Week), shape = 5,data = window1_predict) +
  geom_line(aes(x = Date, y = I,color = Week),data = window1_predict) +
  theme_bw() +
  ylab("Number of New Case (log transformation)") +
  ggtitle("Forecasting for one week window")
```

From the above plot, the forecasting value for period 3 are not very good.

```{r}
predict_df <- tibble("Date" = window1_predict$Date,
                     "w1" = window1_predict$I,
                     "w2" = window2_predict$I,
                     "w3" = window3_predict$I,
                     "w4" = window4_predict$I,
                     "w5" = window5_predict$I,
                     "week" = window1_predict$Week)
predict_df %>%
  pivot_longer(cols = -c("Date", "week"), names_to = "Windows",values_to = "I") %>%
  mutate(type = paste0(week, Windows)) -> predict_long
```

The average for the forecasting value of all window

```{r}
(window1_predict$I + window2_predict$I + window3_predict$I + window4_predict$I + window5_predict$I)/5 -> mean_I

tibble("Date" = window1_predict$Date, "I" = mean_I, "Week" = window1_predict$Week) -> mean_predict
```


```{r}
decomp_comp %>%
  filter(Date >= "2022-07-01") %>%
  ggplot() +
  geom_line(aes(x = Date, y = `log(I)`,color = "Actual"),color = "grey") +
  geom_line(aes(x = Date, y = I,group = interaction(week, Windows), 
                color = Windows) ,data = predict_long)+
  ggtitle("Forecasting value") +
  ylab("Number of New Case (log transformation)") +
  theme_bw()
```

```{r}
decomp_comp %>%
  filter(Date >= "2022-07-01") %>%
  ggplot() +
  geom_line(aes(x = Date, y = `log(I)`,color = "Actual"),color = "grey") +
  geom_line(aes(x = Date, y = I,color = "mean", group = Week), data = mean_predict) +
  geom_point(aes(x = Date, y = I,color = "mean"),shape = 2,data = mean_predict, size = 0.5) +
  theme_bw() +
  ylab("Number of New Case (log transformation)") +
  ggtitle("The Mean of Forecasting Value")
```

The mean value may be better

# Back transformation

```{r}
full_data %>%
  filter(Date >= "2022-07-01") %>%
  ggplot() +
  geom_line(aes(x = Date, y = I,color = "Actual"),color = "grey") +
  geom_point(aes(x = Date, y = exp(I),color = Week), shape = 5,data = window1_predict) +
  geom_line(aes(x = Date, y = exp(I), color = Week),data = window1_predict) +
  theme_bw() +
  ylab("Number of New Case") +
  ggtitle("Forecasting for one week window")
```

```{r}
full_data %>%
  filter(Date >= "2022-07-01") %>%
  ggplot() +
  geom_line(aes(x = Date, y = I,color = "Actual"),color = "grey") +
  geom_line(aes(x = Date, y = exp(I),group = interaction(week, Windows), 
                color = Windows) ,data = predict_long)+
  ggtitle("Forecasting value") +
  ylab("Number of New Case") +
  theme_bw()
```

```{r}
full_data %>%
  filter(Date >= "2022-07-01") %>%
  ggplot() +
  geom_line(aes(x = Date, y = I,color = "Actual"),color = "grey") +
  geom_line(aes(x = Date, y = exp(I),color = "mean", group = Week), data = mean_predict) +
  geom_point(aes(x = Date, y = exp(I),color = "mean"),shape = 2,
             data = mean_predict, size = 0.5) +
  theme_bw() +
  ylab("Number of New Case") +
  ggtitle("The Mean of Forecasting Value")
```

# Using SI mean 7, std 1

```{r}
res_w1 <- r_estiamte(df = data, start = t_one, end = te_one, mean = 20, std = 1)
res_w2 <- r_estiamte(df = data, start = t_two, end = te_two, mean = 20, std = 1)
res_w3 <- r_estiamte(df = data, start = t_three, end = te_three, mean = 20, std = 1)
res_w4 <- r_estiamte(df = data, start = t_four, end = te_four, mean = 20, std = 1)
res_w5 <- r_estiamte(df = data, start = t_five, end = te_five, mean = 20, std = 1)
```

```{r}
tibble("Date" = data$Date[te_one],"R" = res_w1$R$`Mean(R)`, "window" = "w1") -> r_w1
tibble("Date" = data$Date[te_two],"R" = res_w2$R$`Mean(R)`, "window" = "w2") -> r_w2
tibble("Date" = data$Date[te_three],"R" = res_w3$R$`Mean(R)`, "window" = "w3") -> r_w3
tibble("Date" = data$Date[te_four],"R" = res_w4$R$`Mean(R)`, "window" = "w4") -> r_w4
tibble("Date" = data$Date[te_five],"R" = res_w5$R$`Mean(R)`, "window" = "w5") -> r_w5
```

```{r}
bind_rows(r_w1, r_w2) %>%
  bind_rows(r_w3) %>%
  bind_rows(r_w4) %>%
  bind_rows(r_w5) -> df_f
```

```{r}
df_f %>%
  ggplot() +
  geom_line(aes(x = Date, y = R, color = window)) +
  geom_hline(aes(yintercept = 1), color = "grey") +
  theme_bw()
```

```{r}
window_w1 <- find_r(date = range, r_df = res_w1, full_df = full_data_adj)
window_w2 <- find_r(date = range, r_df = res_w2, full_df = full_data_adj)
window_w3 <- find_r(date = range, r_df = res_w3, full_df = full_data_adj)
window_w4 <- find_r(date = range, r_df = res_w4, full_df = full_data_adj)
window_w5 <- find_r(date = range, r_df = res_w5, full_df = full_data_adj)
window1_predict <- forecast_i(r_date = window_w1, full_df = full_data_adj, r_df = res_w1)
window2_predict <- forecast_i(r_date = window_w2, full_df = full_data_adj, r_df = res_w2)
window3_predict <- forecast_i(r_date = window_w3, full_df = full_data_adj, r_df = res_w3)
window4_predict <- forecast_i(r_date = window_w4, full_df = full_data_adj, r_df = res_w4)
window5_predict <- forecast_i(r_date = window_w5, full_df = full_data_adj, r_df = res_w5)
window1_predict <- add_season_pattern(f_data = window1_predict, seasona_data = decomp_comp)
window2_predict <- add_season_pattern(f_data = window2_predict, seasona_data = decomp_comp)
window3_predict <- add_season_pattern(f_data = window3_predict, seasona_data = decomp_comp)
window4_predict <- add_season_pattern(f_data = window4_predict, seasona_data = decomp_comp)
window5_predict <- add_season_pattern(f_data = window5_predict, seasona_data = decomp_comp)
```

```{r}
decomp_comp %>%
  filter(Date >= "2022-07-01") %>%
  ggplot() +
  geom_line(aes(x = Date, y = `log(I)`,color = "Actual"),color = "grey") +
  geom_point(aes(x = Date, y = I, color = Week), shape = 5,data = window1_predict) +
  geom_line(aes(x = Date, y = I,color = Week),data = window1_predict) +
  theme_bw() +
  ylab("Number of New Case (log transformation)") +
  ggtitle("Forecasting for one week window")
```

There are some non-linear trend, The forecasting value for the late time, which is 2022/08/15 to 2022/09/01, is better than model with mean 4.7, and std 2.9

```{r}
predict_df <- tibble("Date" = window1_predict$Date,
                     "w1" = window1_predict$I,
                     "w2" = window2_predict$I,
                     "w3" = window3_predict$I,
                     "w4" = window4_predict$I,
                     "w5" = window5_predict$I,
                     "week" = window1_predict$Week)
predict_df %>%
  pivot_longer(cols = -c("Date", "week"), names_to = "Windows",values_to = "I") %>%
  mutate(type = paste0(week, Windows)) -> predict_long
```

The average for the forecasting value of all window

```{r}
(window1_predict$I + window2_predict$I + window3_predict$I + window4_predict$I + window5_predict$I)/5 -> mean_I

tibble("Date" = window1_predict$Date, "I" = mean_I, "Week" = window1_predict$Week) -> mean_predict
```


```{r}
decomp_comp %>%
  filter(Date >= "2022-07-01") %>%
  ggplot() +
  geom_line(aes(x = Date, y = `log(I)`,color = "Actual"),color = "gray") +
  geom_line(aes(x = Date, y = I, group = interaction(week, Windows), color = Windows) ,
            data = predict_long)+
  ggtitle("Forecasting value") +
  ylab("Number of New Case (log transformation)") +
  theme_bw()
```

```{r}
decomp_comp %>%
  filter(Date >= "2022-07-01") %>%
  ggplot() +
  geom_line(aes(x = Date, y = `log(I)`,color = "Actual"),color = "grey") +
  geom_line(aes(x = Date, y = I,color = "mean", group = Week), data = mean_predict) +
  geom_point(aes(x = Date, y = I,color = "mean"),shape = 2,data = mean_predict, size = 0.5) +
  ggtitle("The mean of Forecasting value") +
  ylab("Number of New Case") +
  theme_bw()
```

Back transformation
 
```{r}
full_data %>%
  filter(Date >= "2022-07-01") %>%
  ggplot() +
  geom_line(aes(x = Date, y = I,color = "Actual"),color = "grey") +
  geom_point(aes(x = Date, y = exp(I), color = Week), shape = 5,data = window1_predict) +
  geom_line(aes(x = Date, y = exp(I),color = Week),data = window1_predict) +
  theme_bw() +
  ylab("Number of New Case (log transformation)") +
  ggtitle("Forecasting for one week window")
```

```{r}
full_data %>%
  filter(Date >= "2022-07-01") %>%
  ggplot() +
  geom_line(aes(x = Date, y = I,color = "Actual"),color = "gray") +
  geom_line(aes(x = Date, y = exp(I), group = interaction(week, Windows), color = Windows) ,
            data = predict_long)+
  ggtitle("Forecasting value") +
  ylab("Number of New Case") +
  theme_bw()
```

```{r}
full_data %>%
  filter(Date >= "2022-07-01") %>%
  ggplot() +
  geom_line(aes(x = Date, y = I,color = "Actual"),color = "grey") +
  geom_line(aes(x = Date, y = exp(I),color = "mean", group = Week), data = mean_predict) +
  geom_point(aes(x = Date, y = exp(I),color = "mean"),
             shape = 2,data = mean_predict, size = 0.5) +
  ylab("Number of New Case") +
  ggtitle("The Mean of Forecasting Value") +
  theme_bw()
```

There is a large difference between the forecasting value and actual value at the end of July.

```{r}
full_data %>%
  filter(Date >= "2022-07-01") %>%
  ggplot() +
  geom_line(aes(x = Date, y = I,color = "Actual"),color = "grey") +
  geom_line(aes(x = Date, y = exp(I),color = "mean", group = Week), data = mean_predict) +
  geom_point(aes(x = Date, y = exp(I),color = "mean"),shape = 2,
             data = mean_predict, size = 0.5) +
  geom_ribbon(aes(Date, ymin = Low, ymax = High, group = Week),data = mean_predict) +
  theme_bw() +
  ylab("Number of New Case") +
  ggtitle("The Mean of Forecasting Value")

mean_predict %>%
  mutate(I = exp(I)) -> mean_predict_exp
  ggplot() +
  geom_line(aes(x = Date, y = exp(I))) +
  geom_ribbon(aes(Date, ymin = Low, ymax = High, group = Week),alpha=0.05, fill = "blue", 
              color = "blue", linetype = "dotted")

```

